Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-13 paper/paper.tex
minor

An implementation of Anglican must therefore address three issues:
\begin{itemize}
    \item the Clojure syntax to introduce probabilistic Anglican
        code inside Clojure modules;
    \item source-to-source transformation of Anglican programs
        into Clojure, so that probabilistic execution becomes
        possible;
    \item algorithms which run Clojure code, obtained by
        transforming Anglican programs, according to the
        probabilistic operational semantics.
\end{itemize}
Execution of probabilistic programs by inference algorithms is
different from execution of deterministic programs.  A
probabilistic program is executed multiple times, often hundreds
of thousands or even millions of times for a single inference
task. Random choices may affect which parts of the program code
are executed and how often. Many inference algorithms require
re-running the program multiple times partially, from a certain
point on.  Different executions may employ different random
choices. However, for efficient inference a correspondence
between random choices in different executions should be
maintained. These are just some of the challenges which were
faced and solved during development of Anglican.

>>>>

Inference algorithms execute Anglican programs while trying to
answer probabilistic queries on those programs. This execution is 
significantly different from the one described in the standard 
operational semantics. These algorithms typically run Anglican programs 
multiple times, often hundreds of thousands or even millions of times 
for a single inference task. The algorithms may make random choices 
that do not correspond to any statements in the program, and decide which parts 
of the program code are executed and how often. Some inference algorithms 
re-run the program multiple times partially, from a certain
point on, while reusing random choices made in the previous runs as much as possible. 
A good high-level picture is that each inference algorithm specifies a 
virtual machine that executes Anglican programs according to a 
non-standard (often probabilistic) operational semantics. Supporting a wide range 
of inference algorithms and their unusual operational semantics efficiently is the main reason 
that we developed Anglican as a language rather than as a library.

%An implementation of Anglican must therefore address three issues:
%\begin{itemize}
%    \item the Clojure syntax to introduce probabilistic Anglican
%        code inside Clojure modules;
%    \item source-to-source transformation of Anglican programs
%        into Clojure, so that probabilistic execution becomes
%        possible;
%    \item algorithms which run Clojure code, obtained by
%        transforming Anglican programs, according to the
%        probabilistic operational semantics.
%\end{itemize}
%Execution of probabilistic programs by inference algorithms is
%different from execution of deterministic programs.  A
%probabilistic program is executed multiple times, often hundreds
%of thousands or even millions of times for a single inference
%task. Random choices may affect which parts of the program code
%are executed and how often. Many inference algorithms require
%re-running the program multiple times partially, from a certain
%point on.  Different executions may employ different random
%choices. However, for efficient inference a correspondence
%between random choices in different executions should be
%maintained. These are just some of the challenges which were
%faced and solved during development of Anglican.
